---
title: 'Coursera Project Practical Machine Learning: Prediction Assignment Writeup'
author: "Thorsten"
date: '2020-01-09'
output:
  pdf_document: default
  html_document: default
---

<!-- Check for dependencies -->
```{r echo=FALSE, warning=FALSE, message=FALSE}
require(caret)
require(randomForest)
require(lattice)
require(ggplot2)
require(datasets)
require(dplyr)
require(doParallel)
```

### Overview
The goal of the project is to predict the manner in which they did the exercise.
Following data will be available:

The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


Applying the machine learning algorithm to the 20 test cases available in the test data for checking the model.

#### Loading data
```{r cache=TRUE}
trainset <- read.csv("pml-training.csv")
```

#### Cleaning data
```{r cache=TRUE}
trainset <- trainset[ , colSums(is.na(trainset)) == 0] # selecting only columns that do not have NAs
trainset <- trainset[ , -nearZeroVar(trainset)] # removing columns with near zero variance
trainset <- trainset[ , -c(1:6)] # removing variables for row number, username, timestamp, numwindow
```

#### Devide trainset into train/test for Prediction

```{r cache=TRUE}
partition <- createDataPartition(y=trainset$classe, p=0.8, list=FALSE)
trainset.Train <- trainset[partition,]
trainset.Test <- trainset[-partition,]
```

### Prediction

##### Parallel Processing
```{r}
cl <- makePSOCKcluster(3)  # use three cores
registerDoParallel(cl)  # do not forget to deregister via stopCluster(cl)
```

```{r}
theControl <- trainControl(method = "repeatedcv", number = 4, repeats = 2, allowParallel = TRUE, verboseIter = TRUE) #cross-validation, 4 folds, 3repeats

```
#### Random Forest

```{r message = FALSE}
theModel <- train(classe ~ ., data = trainset.Train, method = "rf", trControl = theControl)
plot(theModel)
```

##### Stop Parallel Processing
```{r}
stopCluster(cl)
```

```{r}
thePredict <- predict(theModel, trainset.Test)
theConfMat <- confusionMatrix(thePredict, trainset.Test$classe)
theConfMat

plot(theConfMat$table, col = theConfMat$byClass, main = paste("RF - Overall Accuracy = ", round(theConfMat$overall['Accuracy']*100, 2), "%", sep=""))

```

### Prediction on test dataset

```{r}
testset <- read.csv("pml-testing.csv")
testset <- testset[ , colSums(is.na(testset)) == 0]
testset <- testset[ , -nearZeroVar(testset)]

thePredictResult <- predict(theModel, testset)
thePredictResult
```
